An Automatic Verification Framework for Software Configurations

### 2016-05-09 OSDI/arch.tex
config: updated

\para{Checking.}
The checker is used to detect rule violations in the configuration
files of interest. The inputs of the checker are the learned rules 
and the target configuration file to verify.
It generates a report (as shown in $\S$\ref{sec-motiv}) about 
whether it finds any error, \eg, rule violations 
or some suspicious values.
As shown in Figure~\ref{fig-overview},
there are two sub-modules in the checker. They are responsible for
checking rule violations and suspicious values, respectively.
In our experience, we find learned rules could be significantly reused
to check different configuration files, thus improving our usability.

>>>>

\para{Checking.}
The checker is used to detect rule violations in the configuration
files of interest. The inputs of the checker are the learned rules 
and the target configuration file to verify.
It generates a report (as shown in $\S$\ref{sec-motiv}) about 
whether it finds any error, \eg, rule violations 
or some suspicious values (\eg, singular value anomaly).
As shown in Figure~\ref{fig-overview},
there are two sub-modules in the checker. They are responsible for
checking rule violations and suspicious values, respectively.
In our experience, we find learned rules could be significantly reused
to check different configuration files, thus improving our usability.

### 2016-05-09 OSDI/learner.tex
config: updated

\ennan{This section details how do we extract rules from the
intermediate representations. This section may should have the following
two subsections: Aaron's part (including Xinyu's fine-grained
value correlation rule learning) and Jonathon's part.
In addition, we may need to add one more subsection about
learning system environment information related rules,
\eg, for paths and users.}

In general, the learner module has two parts of learning tasks.
One is to learn rules in order to check configuration errors like
entry missing, ordering errors, and correlation errors.
Another one is to learn constraints to check suspicious (or called
anomalous) singular values that may violate standard values.

\subsection{Learning Rules}
\label{subsec-rules}

\subsection{Learning Suspicious Constraints}
\label{subsec-constraints}
Anomalous values can cause tricky performance and memory issues that are hard to debug,
as discussed in Example 4 of Section \ref{sec-motiv}. Consequently, suspicious values 
should be flagged and a warning returned to the user indicating the violation.

We now describe the technique we use to detect anomalous values for 
numerical attributes. Let $A$ be the set of attributes contained in the 
configuration files in the sample dataset. 
Let $A_n$ be the subset of attributes of $A$ which are numerically typed. 
Then, for each attribute $a \in A_n$, we construct a vector $v_a$ of the 
values corresponding to attribute $a$, seen over the entire sample dataset.
For each $v_a$, we compute 
an interval  $$[\hat{v_a} - 3*MAD(v_a), \hat{v_a} + 3*MAD(v_a)],$$ 
where $\hat{v_a}$ represents the median over the values 
in $v_a$ and $MAD(v_a$) refers to the 
median absolute deviation. 
This is a standard outlier detection test, namely the Hampel identifier. 
\footnote{Mathematically, $MAD(v_a) = 1.4826* median(|v_a - \hat{v_a}|)$, estimating standard deviation 
for a normal distribution.} 
In the checking phase, as long as the checker finds a value for a numerical 
attribute in the checked file outside of this interval, 
a warning would be printed to the user indicating the violating value, 
the attribute, and the upper or lower Hampel threshold.

>>>>

The goal of learner module is to derive rules and constraints from
the intermediate representation generated by the translator.
In general, the learner module has two parts.
The first part is to learn rules to check configuration errors like
missing entry, ordering errors, and fine-grained value correlation errors. 
These errors tend to cause total system failures.
%Once the configuration file has been validated against such rules, 
%the user may choose to invoke a more sensitive constraint checker. 
The second part of the learner aims to derive 
constraints on entries to check for suspicious (\ie, anomalous) values 
that may violate standard practice (like the example~4 in
$\S$\ref{sec-motiv}).

\subsection{Derivation of Probabilistic Rules}
\label{subsec-rules}

\para{A strawman solution.}
The first step of \app is to learn relations 
that must hold over multiple parts of a configuration file.
We now first present a strawman solution (employed by
previous work~\cite{santolucitoCAV, zhang14encore}) that uses 
a set of correct configuration files as a learning set, 
from which it was possible to derive rules 
that must hold with absolute certainty. 
However, this is approach is severely limiting. 
It is difficult to obtain a set of files 
that is both guaranteed to be without misconfiguration 
and large enough to learn the many rules of configuration files.
This usually requires manual verification of the learning set, 
which is prone to error.

As a result of this restriction, 
these efforts only consider a rule if it holds over exactly every file in 
the learning set. This behavior can be formally described as follows:

\begin{small}
\begin{flalign*}
C =&\ \text{Correct Learning Set}\\
\text{::}\ & \text{\{Configuration Files in Intermediate Representation\}}\\
LR =&\ \text{Learned Rules :: \{Rule\}}\\
RR =&\ \text{Reported Rules :: \{Rule\}}\\ 
LR =&\ \{ r\ \mid \forall file \in C,\ holds(r,file)\} \\
RR =&\ \{ r\ \mid r \in L\ \land \neg\ holds(r,userfile) \}
\end{flalign*}
\end{small}

Each rule can be thought of as a mapping from 
lines $j$ and $k$ in a configuration file to a relation, $R$.

\[
\{ Rule = (a_j, a_k) | j \neq k \} \rightarrow \{ R \}
\]

Specifically, $a_j$ and $a_k$ are two different lines from our
intermediate representation, or more formally, $j \neq k \land
a_j, a_k \in \{ L \}$ where is the set of lines (key-value pairs) found
in the learning set. The relation $R$
is a Boolean function specific to the error to detect. 
As an example, to detect the error that key $k_1$ must always 
have a value greater than key $k_2$, the relation $R$ is $>$.

\para{Our approach.}
In \app, the rule learning mechanism is tolerant 
enough to accept a dataset {\em full of} incorrect configuration files.
Rather than manually correcting each file, 
%we observe that the files are usually incorrect in several lines.
We extend the previous formalism to run probabilistic learning
on our intermediate representations (generated by the translator).

Our probabilistic approaches for learning the missing values, 
ordering, and fine-grained value correlation rules stem 
from our previous work with building 
the non-probabilistic versions of these rule-learning algorithms. 
We start with the idea that for each of these rules, 
we are going to consider all possible pairs of keys that appear in every 
file, and for our learning process, calculate the likelihood that each of 
these pairs constitute a rule. 
In essence, this is a mechanism that will consider many different possible 
pairs of lines in the file, 
and attempt to compile a set of these pairs 
that are expressed more often than others as a basis for finding patterns 
within the example set that can be used to evaluate new files.

More formally, this approach can be defined as follows. 
The output of our learning algorithm is augmented to a map from 
key-value pairs to a probability distribution over number of possible 
relations.

\[
\{ P\_Rule = (a_j, a_k) | j \neq k \} \rightarrow \{ (R_1, R_2, ... , R_n) \}
\]

We can think of each of these $(a_j, a_k)$ as possibly having a different relationship, defined by the set $\{ R_i \}$, which cover the entire outcome space of possible relationships between the two values $(a_j, a_k)$.

For the entry missing rules, we define $R_1$ as the event that $a_j$ and
$a_k$ appear together, and $R_2$ to be the event that $a_j$ appears
without $a_k$, or by the transitive equivalent, $a_k$ appears without
$a_j$. For the ordering rules, we define $R_1$ as the event that
$a_j$ appears before $a_k$ and $R_2$ be the event that $a_k$ appears
before $a_j$. For the value correlation rules, we define $R_1$ as the
event that $a_j \leq a_k$, $R_2$ the event that $a_j = a_k$, and $R_3$
the case that $a_j \geq a_k$. Notice that the $R_i$ do not have to be
disjoint, but only have to union to the entire probability space.

By examining the learning set, we will derive a distribution of the set $\{R_i\}$ based on how many times we observe an occurrence of each relation. This distribution will then be used at checking time to determine if a user's configuration has broken a likely rule.

\begin{small}
\begin{flalign*}
I\ \ =&\ \text{Incorrect Learning Set}\\
\text{::}\ & \text{\{Configuration Files in Intermediate Representation\}}\\
LP =&\ \text{Learned Probabilistic Rules :: \{(P\_Rule)\}}\\
RP =&\ \text{Reported Probabilistic Rules :: \{(P\_Rule)\}}\\
LP =&\ \text{count\_relation\_occurrences}(I)\\
RP =&\ \{ r\ \mid r \in LP\ \land \Pi(r)>p \land \neg r(userfile) \}
\end{flalign*}
\end{small}

A rule will be reported as broken if the probability the rule is correct, $\Pi$, is greater than some user defined constant, $p$. This constant can be adjust to the user's preference. A small $p$ will increase the likelihood of finding an error, but also increase the number of false positives that are reported.

\subsection{Learning Suspicious Constraints}
\label{subsec-constraints}

With a configuration file that has been verified against catastrophic
failures, the user may also use \app to find more subtle issues.
Anomalous values can cause tricky, but impactful, performance and memory
issues that are hard to debug, as discussed in Example 4 of 
$\S$\ref{sec-motiv}. 
Consequently, suspicious values should be flagged and a warning returned
to the user indicating the violation.

We now describe the technique we use to detect anomalous values for 
numerical attributes. Let $A$ be the set of attributes contained in the 
configuration files in the sample dataset. 
Let $A_n$ be the subset of attributes of $A$ which are numerically typed. 
Then, for each attribute $a \in A_n$, we construct a vector $v_a$ of the 
values corresponding to attribute $a$, seen over the entire sample dataset.
For each $v_a$, we compute 
an interval  $$[\hat{v_a} - 50*MAD(v_a), \hat{v_a} + 50*MAD(v_a)],$$ 
where $\hat{v_a}$ represents the median over the values 
in $v_a$ and $MAD(v_a$) refers to the 
median absolute deviation. 
This is a variant of a standard outlier detection test, namely the Hampel identifier.\footnote{Mathematically, $MAD(v_a) = 1.4826* median(|v_a - \hat{v_a}|)$, estimating standard deviation 
for a normal distribution.} 
In the checking phase, as long as the checker finds a value for a numerical 
attribute in the checked file outside of this interval, 
a warning would be printed to the user indicating the violating value, 
the attribute, and the upper or lower Hampel threshold.

### 2016-05-09 OSDI/trans.tex
config: checker part
config: updated

\section{Translator}

\ennan{Translator is, in some sense, like a parser. 
This section is used to detail how we design the parser.}

The responsibility of the translator is to transform a given set of
configuration files into another set of files written in
typed and well-structured representations.
In some sense, the translator looks like a parser, and it is only
used to generate an intermediate representation for the post
process, such as learning rules (see $\S$\ref{sec-learn}).
Coupled with rules generated by learner module,
we will have a complete language model to verify configuration files
of our interest.

Translating or parsing is system dependent. In other words, for MySQL
and HDFS, we need to develop different parsers to handle each of them,
respectively. \app allows users to provide extra help to the translator
for their specific system configuration, such as specifying a comment 
character. Each of the entries in the intermediate representations
must be assigned types, and these types would guide the learning
modules later.

@@

\para{Introducing probabilistic types.}
Based on a single example value of $v$,
the translator cannot fully determine the type of $k$.
Consider for instance the following example:

@@

Note that typing is also a system module than 
can be easily extended to support more types. 
In that case the user will need to provide rules for type inference 
and probability distributions for values where type inference is ambiguous.

>>>>

\section{Translator}
\label{sec-trans}

The responsibility of the translator is to transform a given sample
dataset of configuration files into another set of files written in
typed and well-structured representations.
In some sense, the translator looks like a parser, and it is only
used to generate an intermediate representation for the post
process, such as learning rules (see $\S$\ref{sec-learn}).
Coupled with rules generated by learner module,
we will have a complete language model to verify configuration files
of our interest.

Translating or parsing is system dependent. In other words, for MySQL
and HDFS, we need to develop different parsers to handle each of them,
respectively. \app also allows users to 
provide extra help to the translator
for their specific system configurations,
but not necessarily. 
Each of the entries in the intermediate representations
have been assigned types, and these types would guide the learning
modules later.

@@

\para{Introducing probabilistic types.}
For any key-value entry in the sample configuration files,
the translator cannot fully determine the type of key 
based on a single example value.
Consider the following example.

@@

\com{Note that typing is also a system module than 
can be easily extended to support more types. 
In that case the user will need to provide rules for type inference 
and probability distributions for values where type inference is ambiguous.
}
